---
title: "Assignment 5"
output: pdf_document
---
  
```{r, echo = FALSE, include=FALSE}

##### Load libraries ######
library(tidyverse)
library(fpp3)
library(kableExtra)
library(fGarch)
library(lubridate)
library(readxl)
library(magrittr)
library(forecast)
library(dynlm)
library(fable)
library(tseries)

```



```{r, echo = FALSE, include=FALSE}

### Data retrieval ##

ets = read_csv("http://jmaurit.github.io/analytics/labs/data/eua-price.csv") ## Carbon pricing data
colnames(ets) = c("date", "price")                                                # Change rownames
elspot = read_csv("http://jmaurit.github.io/norwayeconomy/data_series/elspot.csv") ##


ets["month"] = month(ets$date)
ets["year"] = year(ets$date)

ets_mon = ets %>% group_by(month, year) %>% summarise(
  price = mean(price))

ets_mon["day"] = 1

ets_mon = ets_mon %>% arrange(year, month)

ets_mon = ets_mon %>% mutate(date = make_date(year, month, day))

## Convert to monthly data
ets %<>% 
  mutate(month = month(date)) %>% 
  group_by(month) %>% 
  summarise(ets_mon = mean(price),
            year = year(date),
            month = month,
            day  = 1) %>%  
  ungroup() %>% 
  mutate(date = make_date(year, month, day ))


# Join carbon pricing data and 
power_df <- elspot %>% inner_join(ets_mon[c("price", "date")], by="date")
power_DK_df <- power_df %>% dplyr:: select(DK1, DK2, date, price)


## Scale to mwh
power_DK_df %<>% mutate(DK1 = DK1/1000,
                        DK2 = DK2/1000)


## Daily consumption data
cons = read_csv2("http://jmaurit.github.io/analytics/labs/data/consumption-per-country_2019_daily.csv")
cons["date"] = as.Date(cons$date, format="%d/%m/%Y")
cons_ts <- tsibble(cons, index=date)


```


## Task 1: Comparison of consumption data in Norway and Denmark


We start our comparison of norwegian and danish power consumption by looking at the STL decompositon plots.


```{r, echo = TRUE}
## Seasonal decompositon of Danish electricity consumption

cons_comp_dk = cons_ts %>% model(
  STL(DK ~ trend(window=7) + season(window="periodic"))
) %>% components 

cons_comp_no = cons_ts %>% model(
  STL(NO ~ trend(window=7) + season(window="periodic"))
) %>% components 


cons_comp_dk %>%  autoplot()

cons_comp_no %>%  autoplot()

```
From the components we can see the same general trend of winter increase, and corresponding the decrease in power consumption in the summer months.
Where they differ is the magnitude of the weekly-sesaonal component. In Denmark the weekend effect of reduced power consumption is more noticable.

Another small difference is the effect of winter, where it appears that the winter-increase in power consumption is larger in Norway than Denmark.

### Forecasting using a seasonal ARIMA model

A clear assumption in an ARIMA forecasting model is the that the data is stationary in terms of its variance and mean. We plot the time series containing electricity consumption data, as well its autocorrelation and partial autocorrelation plots. 
We see clear signs of non-stationarity and perform unit root tests confirming the need for differencing. 
As there appears to be a strong seasonal autocorrelation, we will first conduct a seasonal differencing, and see if this solves our non-stationarity issue.
If not, further differencing will be needed. 
Some information contained in the data is lost by performing a differencing, but we conform the the assumption of stationarity of the data.


```{r, echo = TRUE, include = TRUE}
forecast::ggtsdisplay(cons$DK, plot_type='partial',
                      lag.max = 24, 
                      theme = theme_bw(),
                      main = "Elecitricity consumption in Denmark ACF and PACF plots ") 

unitroot_kpss(cons$DK)
adf.test(cons$DK)

```


We perform a first order differencing and perform the same stationarity analysis.
```{r, echo = TRUE, include = TRUE}

# Perform differencing

cons_diff_dk <- cons %>% mutate(DK = difference(DK,7)) %>% dplyr::filter(!is.na(DK)) #Take first order difference

unitroot_kpss(cons_diff_dk$DK)
adf.test(cons_diff_dk$DK) #Stationary


# New ACF and PACF plots

forecast::ggtsdisplay(cons_diff_dk$DK, plot_type='partial',
                      lag.max = 24, 
                      theme = theme_bw(),
                      main = "Elecitricity consumption in Denmark (difference) ACF and PACF plots ") 


```

We note that there are significant autocorrelations at the weekly lag (i.e 7, 14). Luckily the fable package correctly identified the seasonality as weekly, e.g an ARIMA PDQ pdq[7], regardless of the specific terms. 

We will now perform two forecasts, a manually specified ARIMA model and an automatically determined ARIMA model made by the fable ARIMA() function.
In the PACF plot we can see a clear autocorrelation in seasonal lag terms, in a decreasing fashion. This calls for an MA(1) term to applied to the seasonal component of the ARIMA model. 
There is a significant but decreasing correlation at lag 1  as shown in the PACF plot, and an AR(1) term applied to the non-seasonal component is appropiate. 
As the lags are decreasing after 1, and MA(1) term might also be necessary. 

A such we might reason that our model might something like:
      ARIMA pdq(1,0,1) pdq(1,1,1)

```{r, echo = TRUE, include = TRUE}


fit_arima_optimal_cons_dk <- 
  cons %>% 
  as_tsibble(index = date) %>%  
  model(arima_optimal =  ARIMA(DK, stepwise = FALSE, approximation = FALSE))


fit_arima_manual_cons_dk <- cons %>% as_tsibble(index = date) %>%  
  model(arima_101111       = ARIMA(DK ~ 0 + pdq(1,0,1) + PDQ(1,1,1)))

fit_cons_dk  <-  fit_arima_manual_cons_dk  %>% bind_cols(fit_arima_optimal_cons_dk)


```

In the plot below we have used our manually selected model to forecast danish power consumption for a time horizon of 30 days.

```{r, echo = TRUE, include = TRUE}
#Forecast for h = 30
fc_cons_dk <-  fit_cons_dk %>% forecast(h = 30)


fc_cons_dk  %>% ggplot() +
  geom_line(aes(x = date, y = .mean, color = .model)) +
  geom_line(aes(x  = date, y = DK), data = cons) +
  theme_bw()   +  
   labs(title = "Forecasting of danish power consumption",
         y = "Consumption", 
         x = "Day") 

```




As we can see our manually selected model slightly outperforms the optimally selected model based on AIC.
The Ljung Box reveals that there is no residual autocorrelation and as such much of the error term is explained in our model.

```{r, echo = TRUE, include = TRUE}

# Forecast evaluation

fit_cons_dk %>%  accuracy()  %>%   
  rename("Model" = .model) %>% 
  kbl(caption = "Training data performance metrics: Danish power consumption") %>%
  kable_classic(full_width = F, html_font = "Times new roman")


# Autocorrelation tests

fit_cons_dk  %>% 
     augment()  %>% dplyr::filter(.model == "arima_101111")
feasts::ljung_box(
  (fit_cons_dk  %>% 
     augment()  %>% dplyr::filter(.model == "arima_101111"))$.innov) # Passing



```
Perhaps the most notable weakness of such a model is its weekly sesaonality. It would be interesting to tweak the model with longer seasonality, as a week may be too small
a period to capture the winter/summer differences in consumption. 

Another weakness is that some information useful for forecasting may be found in other variables, and as a result a multivariate model should also be tested. 





## Task 2: 

We retrieve finnish electricity price data from Nordpool and clean them. 

```{r, echo = TRUE, include = TRUE}

elspot_data <- read_csv("elspot-prices_2019_daily_nok.csv") # Load data

colnames(elspot_data) <- elspot_data[2,]
elspot_data <- elspot_data[3:nrow(elspot_data),]
colnames(elspot_data)[1] <- "date"

elspot_data  %<>% 
    mutate(date = lubridate::dmy(date),
           FI = as.numeric(gsub(",", ".", FI)))  %>%  #Substitute commas
    select(date, FI)  %>% 
    as_tsibble(index = date)

#Plot of finnish electricity prices and ACF/PACF plots
forecast::ggtsdisplay(elspot_data$FI, plot_type='partial',
                      lag.max = 24, 
                      theme = theme_bw(),
                      main = "Finnish electricity prices in 2019 NOK mwh") 

```

We find the conditional variance of the series by calculating the residuals after an ARIMA model (in our case an ARIMA(1,1) model). 
By subtracting the original values with with the fitted values of this ARIMA model, we end up with the "return residuals". 

```{r, echo = TRUE, include = TRUE}

fi_garch <- garchFit(~arma(1,1) + garch(1,1), data = elspot_data$FI, trace = F)
summary(fi_garch)


```

We note the signifiance of the alpha and gamma parameters, and our time series variance appears to be conditional on previous variance.


We model the conditonal volatility of the time series, and note that there are several significant spikes in volatility of electricity prices. For instance around the months May, June, and October. 
```{r, echo = TRUE, include = TRUE}


elspot_data  %<>% mutate(volatility = fi_garch@h.t)
elspot_data  %>% 
    ggplot() + 
    geom_line(aes(y = volatility, x = date))  +
    labs(title = "GARCH modeled volatility: Finnish electricity prices",
         y = "Consumption", 
         x = "Day") +
     scale_colour_manual(values = c("black", "orange"))

```


Using our model we forecast 30 days ahead with prediction intervals. 
```{r, echo = TRUE, include = TRUE}
fc_garch = predict(fi_garch, n.ahead = 30)
fc_garch["date"] = seq(as.Date("2020-01-01"), as.Date("2020-01-01")+29, by = "days")

elspot_data  %>% 
  ggplot() +
  geom_line(aes(x = date, y = FI, color = "Conditional variance")) +
  geom_line(aes(x = date, y = meanForecast, color = "Forecast"), data =  fc_garch) +
  geom_ribbon(aes(x=date, ymin=(meanForecast-2*standardDeviation), ymax=(meanForecast+2*standardDeviation)), data = fc_garch, fill="orange", alpha=.2) +
  scale_colour_manual(values = c("black", "orange")) +
  labs(title = "GARCH modeled volatility: Finnish electricity prices",
         y = "Conditonal variance", 
         x = "Day",
         colour = "Series") 

```



## Task 3

The inclusion of additional independent variables in a forecasting model may be helpful if they provide additional causal information to the model.
The amount of rain has a significant and causal relation to power prices in Norway, as the availability of rain increases the supply of hydroelectric power. 
As such, it  is plausible that such a multivariate model of rain amount and previous power prices may yield a better result than a univariate model based on power prices.
Causality ensures future stability, as previous correlations may disappear going forward in time. 
As causality may be very hard to determine, including a variable may contribute more noise than information for a forecasting model.


The future is uncertain, and addingindependent variables may only be prudent if it contributes significant information.