---
title: "Assignment 5"
output: pdf_document
---

```{r, echo = FALSE, include=FALSE}

##### Load libraries ######


library(tidyverse)
library(fpp3)
library(fGarch)
library(lubridate)
library(magrittr)
library(forecast)
library(dynlm)
library(fable)
library(tseries)

```



```{r, echo = FALSE, include=FALSE}

### Data retrieval ##

ets = read_csv("http://jmaurit.github.io/analytics/labs/data/eua-price.csv") ## Carbon pricing data
colnames(ets) = c("date", "price")                                                  # Change rownames
elspot = read_csv("http://jmaurit.github.io/norwayeconomy/data_series/elspot.csv") ##




ets["month"] = month(ets$date)
ets["year"] = year(ets$date)

ets_mon = ets %>% group_by(month, year) %>% summarise(
  price = mean(price))

ets_mon["day"] = 1

ets_mon = ets_mon %>% arrange(year, month)

ets_mon = ets_mon %>% mutate(date = make_date(year, month, day))




## Convert to monthly data
ets %<>% 
  mutate(month = month(date)) %>% 
  group_by(month) %>% 
  summarise(ets_mon = mean(price),
            year = year(date),
            month = month,
            day  = 1) %>%  
  ungroup() %>% 
  mutate(date = make_date(year, month, day ))
  



# Join carbon pricing data and 
power_df <- elspot %>% inner_join(ets_mon[c("price", "date")], by="date")
power_DK_df <- power_df %>% dplyr:: select(DK1, DK2, date, price)


## Scale to mwh
power_DK_df %<>% mutate(DK1 = DK1/1000,
                        DK2 = DK2/1000)

# Danish power price and carbon pricing
power_DK_df %>% pivot_longer(cols = -date, names_to = "series", values_to = "prices") %>% 
  ggplot() +
  geom_line(aes(x  = date, y = prices, color = series))



####### Stationarity tests ####

dk_power_ts <- power_DK_df %>% mutate(date = yearmonth(date)) %>%  as_tsibble(index = date)

### ACF and PACF plot
forecast::ggtsdisplay(dk_power_ts$DK1, plot_type='partial',
            lag.max = 24, 
            theme = theme_bw(),
            main = "ACF and PACF plot Danish power prices") 

### PACF plot


adf.test(dk_power_ts$DK1)

unitroot_kpss(dk_power_ts$DK1) # Failed unit root test. Needs differencing

dk_power_ts_diff <- dk_power_ts %>% mutate(DK1 = difference(DK1)) %>% filter(!is.na(DK1)) %>% as_tsibble(index = date)

unitroot_kpss(dk_power_ts_diff$DK1) # Passes. No further differencing

## Build optimal ARIMA model



dk_power_ts_diff %>% fill_gaps() %>%  model(arima = ARIMA(DK1 ~pdq(0,0,0)))


fit_dk_power_arima <- dk_power_ts_diff %>% fill_gaps() %>%   
  model(arima_optimal = ARIMA(DK1, 
                              stepwise = FALSE, 
                              approximation = FALSE))


fc_dk_power_arima <- fit_dk_power_arima %>% forecast(h = 10 )


## Daily consumption data
cons = read_csv2("http://jmaurit.github.io/analytics/labs/data/consumption-per-country_2019_daily.csv")
cons["date"] = as.Date(cons$date, format="%d/%m/%Y")
cons_ts = tsibble(cons, index=date)
cons_ts %>% autoplot(NO)


## Decomposition using STL method
cons_comp = cons_ts %>% model(
  STL(NO ~ trend(window=7) + season(window="periodic"))
) %>% components 

cons_comp %>% autoplot()

cons_ts["NO_sa"] = cons_comp$season_adjust
cons_ts = tsibble(cons_ts, index=date)



cons_ts %>%
  gg_tsdisplay(difference(NO, 7), plot_type='partial') 


sfit1 <- cons_ts %>%
  model(
    arima101011 = ARIMA(NO ~ 0 + pdq(1,0,1) + PDQ(0,1,1)),
    auto = ARIMA(NO)
  )

sfit1 %>% select(arima101011)%>%  report()
sfit1 %>% select(auto)%>%  report()
sfit1 %>%  select(auto) %>%  forecast(h=120) %>% autoplot(cons_ts)
cons_ts =  cons_ts %>% mutate(
   returns = log(NO) %>% difference(7)
 )

 
cons_ts %>% autoplot(returns)
cons_ts = cons_ts %>% filter(!is.na(returns))

arch_mod1 = cons_ts %>%
  model(
    arima100100 = ARIMA(returns ~ 0 + pdq(1,0,0) + PDQ(1,0,0))
  )

arch_mod1 %>% report()
#create the
resids = arch_mod1 %>% residuals()

resids = resids %>% mutate(
  res_sq = .resid^2
)

#modell for squared-residuals
arch_reg = resids %>%
  model(
    arima100 = ARIMA(res_sq ~ 0 + pdq(1,0,0) + PDQ(0,0,0))
  )

arch_reg %>% report()
cons_ts = cons_ts %>% dplyr::filter(!is.na(returns))
#cons_ts = as_tsibble(cons_ts, index=date)

garchMod1 = garchFit(~arma(1,1) + garch(1,1), data = cons_ts["returns"], trace = F)
summary(garchMod1)
cons_ts["volatility"] =garchMod1@h.t
ggplot(cons_ts, aes(y = volatility, x = date)) + geom_line(col = '#ff9933') + ylab('Conditional Variance') + xlab('Date')
predict1 = predict(garchMod1, 120)
predict1["date"] = seq(as.Date("2020-01-01"), as.Date("2020-01-01")+119, by="days")
cons_fcst = cons_ts %>% full_join(predict1, by="date")

cons_fcst["date"] = as.Date(cons_fcst$date)
cons_fcst %>% 
  ggplot() + 
  geom_line(aes(x=date, y=returns)) +
  geom_line(aes(x=date, y=meanForecast, color="red")) +
  geom_ribbon(aes(x=date, ymin=(meanForecast-2*standardDeviation), ymax=(meanForecast+2*standardDeviation) ), fill="blue", alpha=.2)

```


## Assignment 1: Comparison of variance of consumption data in Norway and Denmark

```{r, echo = TRUE}

## Seasonal decompositon of Danish electricity consumption

cons_comp_dk = cons_ts %>% model(
  STL(DK ~ trend(window=7) + season(window="periodic"))
) %>% components 

cons_comp_dk %>%  autoplot()

```


```{r, echo = TRUE}
forecast::ggtsdisplay(cons$DK, plot_type='partial',
            lag.max = 24, 
            theme = theme_bw(),
            main = "Elecitricity consumption in Denmark ACF and PACF plots ") 


unitroot_kpss(cons$DK)
adf.test(cons$DK)


```
A clear assumption in an ARIMA forecasting model is the that the data is stationary in terms of its variance and mean. We plot the time series containing electricity consumption data, as well its autocorrelation and partial autocorrelation plots. We see clear signs of non-stationarity and perform unit root tests confirming the need for first order differencing. We lose some information contained in the data by performing a differencing, but we conform the the assumption of stationarity of the data.

