---
title: "Assignment 3"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include = FALSE}
#In the majority of real-life analysis situations, you won’t be presented with a complete data set.
#Instead, you will need to create a data set by combining multiple series and sets from different sources. 
#Combining multiple data sets is called Relational Data.

#setwd("C:/Users/Jon/OneDrive/NHH/Master/ENE434/LABS")
rm(list=ls())

#Load libraries:
library(tidyverse)
library(lubridate)

#Load data: 
exploration = read_csv("https://jmaurit.github.io/analytics/labs/data/wellbore_exploration_last_10_years.csv")

exploration["wlbCompletionDate"] = as.Date(exploration$wlbCompletionDate, format = "%d.%m.%Y")
exploration["wlbEntryDate"] =as.Date(exploration$wlbEntryDate, format = "%d.%m.%Y")

sapply(exploration, class)

#Frequency table over the types of wells (Wildcat or appraisal)
exploration %>% 
  group_by(wlbPurpose, wlbContent) %>% 
  summarise(n())

#Visualization of the categories of wildcat drills:
exploration %>% 
  filter(wlbPurpose == "WILDCAT") %>% 
  ggplot + geom_bar(mapping = aes(x=wlbContent))

#Creating a new variable 'year'
exploration["year"] = year(exploration$wlbEntryDate)

#Number of wells drilled each year from 2009 to 2019.
exp_per_year = exploration %>% 
  group_by(year) %>% 
  summarise(n())
colnames(exp_per_year)[2] = "number_wells"
exp_per_year

exploration %>% 
  filter(wlbPurpose == "WILDCAT") %>% 
  ggplot + geom_bar(mapping = aes(x=year))


#Want to combine/merge the drilling data with financial data
#Financial data from Proff Forvalt
AS_data = read.csv("https://jmaurit.github.io/analytics/labs/data/accounting_cleanData.csv", stringsAsFactors=FALSE)
AS_data["year"] = as.numeric(AS_data$year)

company = read_csv("https://jmaurit.github.io/analytics/labs/data/company.csv")
company["Orgnr"] = as.numeric(company$Orgnr)

#Only selecting a few of the columns from the company dataset. 
#Inner joining the company dataset with AS_data. Matching the data with the common columns Orgnr and comp_id
AS_merged = company %>% 
  select(Orgnr, NPD_id, LongName, ShortName) %>% 
  inner_join(AS_data, by = c("Orgnr" = "comp_id"))

#Trying to match the NPD_id with the license in the exploration data
licence = read_csv("http://jmaurit.github.io/analytics/labs/data/licence.csv")

colnames(licence) = c("prlName", "ValidFrom", "ValidTo", "LongName",      
                      "licenseeInterest", "licenseeSdfi",         
                      "operValidFrom", "operDateValidTo",      
                      "NpdidLicence", "cmpNpdidCompany",         
                      "LicenseeDateUpdated", "DatesyncNPD")

licence["ValidFrom"]  = as.Date(licence$ValidFrom, format="%d.%m.%Y")
licence["ValidTo"]  = as.Date(licence$ValidTo, format="%d.%m.%Y")

licence[is.na(licence$ValidTo),"ValidTo"] = as.Date("2020-01-01")


#When we joing the exploration and licence data, we want to match not just on company_id, but also on date. 
#Problem: licence data has two date columns: ValidFrom and ValidTo.

#We solve this with a left join to keep alle the information and matches the exploration data with the license data. 
#We then filter the entries to match our data criteria. 
expl_lic = exploration %>% 
  left_join(licence, by = c("wlbProductionLicence" = "prlName")) %>%
  filter(wlbEntryDate >= ValidFrom, wlbEntryDate < ValidTo)

#Only include the observations where the licence is also the drilling operator. (Imperfect solution)
expl_oper = expl_lic %>% filter(wlbDrillingOperator==LongName)
colnames(expl_oper)[22] = "NPD_id"

#Assignment 1: TASK 1 (Merging financial data (AS_merged) and exploration data (expl_oper))
expl_AS = expl_oper %>% inner_join(AS_merged, by=c("NPD_id", "year"))

compWells = expl_oper %>% group_by(NPD_id, year) %>% summarise(n())
colnames(compWells)[3] = "numbWells"
explAgg_AS = compWells %>% inner_join(AS_merged, by=c("NPD_id", "year"))


explAgg_AS %>% ggplot(aes(x=total_assets,y=numbWells)) +
  geom_point() +
  geom_smooth(method="lm")

explAgg_AS %>% ggplot(aes(x=total_assets,y=numbWells, color=factor(NPD_id))) +
  geom_point(show.legend=FALSE)

#Trying to identify the biggest company (Answer: Equinor)
explAgg_AS %>% filter(total_assets > 3e8) %>% select(name) %>%unique()

#Profitability
explAgg_AS %>% ggplot(aes(x=profitability,y=numbWells, color=factor(NPD_id))) +
  geom_point(show.legend=FALSE)

#Time trends for each company 
explAgg_AS %>% ggplot(aes(x=year, y=numbWells, color=factor(NPD_id))) +
  geom_line(show.legend=FALSE) 
```



# Assignment 3

When we merged the company data (AS_merged) with the exploration data expl_oper, we lost approximately 90 observations. Create a data frame of these 90 observations. Why were they dropped? What do they have in common? Are the loss of these observations problematic for our further analysis? How so?


## TASK 1
```{r}
expl_AS_full_join = expl_oper %>% left_join(AS_merged, by=c("NPD_id", "year"))

diff_set = setdiff(expl_AS_full_join, expl_AS)
```

Questions: Why were they dropped? What do they have in common? Are the loss of these observations problematic for our further analysis? How so?

### Why were they dropped? 

They were dropped because the NPD_id and year values in exploration (expl_oper) data set did not correspond with any similar values in the financial data set (AS_merged).For a left join / merge between two rows to be made both the NPD_id and year much find an equal pairing in the other data set.

### What do they have in common?







## TASK 2


```{r}
library(tidyverse)
setwd("C:/Users/Jon/OneDrive/NHH/Master/V2021/ENE434/LABS/Datafiles/")
brent_crude_set <- read.csv("Europe_Brent_Spot_Price_FOB.csv") %>% 
  tail(.,-4) %>% 
  mutate(year = as.numeric(rownames(.)),
         Europe.Brent.Spot.Price.FOB = as.numeric(Europe.Brent.Spot.Price.FOB)) %>% 
  rename(spot_price = Europe.Brent.Spot.Price.FOB)
```

Compare the average yearly (brent) oil price with expl_AS. Merging by year.

```{r}
explAGG_spot = explAgg_AS %>% 
  left_join(brent_crude_set, by = c("year"))


explAGG_spot %>% group_by(spot_price) %>%
  summarise(total_n_wells = sum(numbWells)) %>%
  ggplot(aes(x= spot_price, y= total_n_wells)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs()



```

## TASK 3


Overview of the CO2 taxes in the period 2009-2020. 
```{r}
CO2_tax = tibble(
  year=2009:2020,
  CO2_tax_NG_krSM3=c(0.49, 0.51, 0.44, 0.45, 0.46, 0.66,.82, 0.84, 0.90, 1, 1, 1.08)
  )

CO2_tax %>% ggplot(aes(x=year, y=CO2_tax_NG_krSM3)) + 
  geom_line()
```

Question: Do you think that this may effect exploratory drilling in the period studied? Do you think a correlation could be estimated with the given data (and with the addition of CO2-tax data)? Why or why not?


## TASK 4

```{r}
reg1 = lm(numbWells ~ total_assets + profitability, data=explAgg_AS)
summary(reg1)
```
Total_assets is more significant than profitability. The R-squared is quite low. 


