---
title: "Lab 7 /Assignment 3 "
output: pdf_document
---
  

```{r, echo = FALSE, include=FALSE}

### Libraries
library(tidyverse)
library(lubridate)
library(magrittr)
library(splines)

pv_df= read_csv("http://jmaurit.github.io/analytics/labs/data/pv_df.csv")

pv_df = pv_df %>% arrange(date) %>% mutate(
  cum_cap = cumsum(nameplate)
)
pv_df %<>% mutate(log2_cum_cap = log2(cum_cap),
                 log2_cost_per_kw = log2(cost_per_kw),
                 monthdate =  as.Date(paste(pv_df$year, pv_df$month, "1",sep = "-")))  %>%  
                 filter(cost_per_kw !=0)

# Save pv_df
#save(pv_df, file = "pv_df.Rdata")
load("pv_df.Rdata")
######### Models from the lab ###############
 

# Linear model
cap_mod = lm(log2_cum_cap ~ monthdate, data=pv_df) 
new_data = tibble(
  monthdate = seq(ymd("2015-01-01"), ymd("2019-12-31"), by="months")
)

new_data["log2_cum_cap"] = predict(cap_mod, newdata=new_data)

learning_mod1 = lm(log2_cost_per_kw~log2_cum_cap, data=pv_df)

#Polynomial model
learning_mod2 = lm(log2_cost_per_kw~poly(log2_cum_cap,4), data = pv_df)

# Splines 
learning_mod3 = lm(log2_cost_per_kw~bs(log2_cum_cap, knots=c(7, 10, 13, 16)), data=pv_df)


```


## Task 1

We interpret this task in the following way:
For two seperate intervals of time (before 2012, and 2012 onwards), make linear learning curves determined
by the fitted values of a simple linear regression of the log base 2 of the cumulative solar capacity and solar price.

Note that we include 2012 in the 'post' data as we found that the dataset stops at 2014. An exclusion of 2012 would have affected the results substantially,
as there would have been too few datapoints to get a proper result.  

Seperating pre and post 2012 data.
```{r, echo = TRUE, include = TRUE, warning = FALSE}
df_pre_2012 = pv_df %>%  filter(year < 2012)
df_post_2012 = pv_df %>% filter(year >= 2012)

```


Creating the variable cum_cap, which calculates the cumulative sum of nameplate.
```{r, echo = TRUE, include = TRUE, warning = FALSE}
df_pre_2012 = df_pre_2012 %>% arrange(date) %>% mutate(
  cum_cap = cumsum(nameplate)
)

df_post_2012 = df_post_2012 %>% arrange(date) %>% mutate(
  cum_cap = cumsum(nameplate)
)
```


Removing zero values
```{r, echo = TRUE, include = TRUE, warning = FALSE}
df_pre_2012 = df_pre_2012 %>% filter(cost_per_kw != 0)
df_post_2012 = df_post_2012 %>% filter(cost_per_kw != 0)

```


Creating the variables *log2_cum_cap* and *log2_cost_per_kw* 

```{r, echo = TRUE, include = TRUE, warning = FALSE}
df_pre_2012["log2_cum_cap"] = log2(df_pre_2012$cum_cap)
df_pre_2012["log2_cost_per_kw"] = log2(df_pre_2012$cost_per_kw)

df_post_2012["log2_cum_cap"] = log2(df_post_2012$cum_cap)
df_post_2012["log2_cost_per_kw"] = log2(df_post_2012$cost_per_kw)
```

Signficance tests show that the capacity has a statistical signifcant correlation with the solar price.
As we will see later, plotting the regression with observed data confirms this.  

```{r, echo = TRUE, include = TRUE, warning = FALSE}
learning_mod_pre = lm(log2_cost_per_kw~log2_cum_cap, data = df_pre_2012)
summary(learning_mod_pre)

learning_mod_post = lm(log2_cost_per_kw~log2_cum_cap, data = df_post_2012)
summary(learning_mod_post)
```


After splitting and calculating the log base 2 of the cumulative solar capacity and solar cost, we plot the fitted 
values of the linear regression with the observed values.

```{r, echo = TRUE, include = TRUE, warning = FALSE}
df_pre_2012  %>% filter(year < 2012)  %>%  ggplot(aes(x = log2_cum_cap, y = log2_cost_per_kw)) +
  geom_point(alpha=.5, color = "black") +
  geom_smooth(method = "lm", color = "orange") +
  labs(title = "Linear learning curve: -2012 data", 
       x     = "Log base 2 cumulative solar capacity",
       y     = "Log base 2 cost per kw electricity generated") +
  theme_bw()


df_post_2012  %>% filter(year > 2012) %>%  ggplot(aes(x = log2_cum_cap, y = log2_cost_per_kw)) +
  geom_point(alpha=.5, color = "black") +
  geom_smooth(method = "lm", color = "orange") +
  labs(title = "Linear learning curve: 2012- data", 
       x     = "Log base 2 cumulative solar capacity",
       y     = "Log base 2 cost per kw electricity generated") +
  theme_bw() 

```
By using simple linear regression we avoid overfitting, as higher-order polynomials might incurr. While, our estimates does not overfit, 
we do in all likelihood suffer from signifcant bias as we assume a linear relationship between the cumulative solar capacity and solar price.
In reality, a linear relationship is very unlikely and as a result a linear model will give poor predictive performance.

We do not however, capture large amount of the information present within the data. 
As we can see from the plots above, there is signifcant variation and this variation will only be accounted for to a small extent. 


## Task 2
For this task we will explore the relationship between cumulative capacity and solar cost using Local Linear Regression.


# Plot fitted values and groups

```{r, echo = TRUE, include = TRUE, warning = FALSE}

pv_df_log <- pv_df %>% 
                mutate(log2_cost_per_kw = log2(cost_per_kw),
                      log2_cum_cap = log2(cum_cap),
                      monthdate =  as.Date(paste(pv_df$year, pv_df$month, "1",sep = "-")))  %>% 
                filter(!is.na(cost_per_kw))  %>% 
                filter(!is.na(monthdate))



pv_df  %>% 
  ggplot(aes(log2_cum_cap, log2_cost_per_kw)) + 
  geom_point() +
  geom_smooth(method = "loess", se = FALSE) +
  labs(title = "Fitted values of Local Regression",
       x = "Cumulative solar capacity",
       y = "Solar cost") +
  theme_bw()

```

Local regression is different from splines in that a set amount of data points is used for each regression.
The regressions are made by a local group of data poins, rather than a set amount of knots. The methods are similar in
that they are a set of regressions conducted on different parts of the dataset.


# Point forecasts
```{r, echo = TRUE, include = TRUE, warning = FALSE}

# Mean solar log cost per month
pv_df_month <- pv_df  %>% group_by(monthdate)  %>% summarise(log2_cost_per_kw = mean(log2_cost_per_kw))  %>% as_tibble()

loess_fit <-   pv_df %>% loess(log2_cost_per_kw~log2_cum_cap, span=.2, data = ., control=loess.control(surface="direct"))

# New dates as tibble
new_data_fc_loess = tibble(
  monthdate = seq(ymd("2015-01-01"), ymd("2019-12-31"), by="months")
)

# Assuming linear estimate of capacity
new_data_fc_loess["log2_cum_cap"] = predict(cap_mod, newdata = new_data_fc_loess)
# Adding local regression predictions of solar cost
new_data_fc_loess["fc_cost"] = predict(loess_fit, newdata = new_data_fc_loess)


# Plotting forecast of cost
new_data_fc_loess   %>% 
   as_tsibble(index = monthdate)  %>% 
   dplyr::select(fc_cost, monthdate)  %>% 
   ggplot() +
   geom_line(aes(x = monthdate, y = fc_cost, col = "Forecast of solar cost")) +
   geom_point(aes(x = monthdate, y = log2_cost_per_kw, color = "Log base 2 solar cost"), data = pv_df_month) +
   theme_bw() +
   labs(x = "Months",
        y = "Solar cost log base 2",
        title = "Forecast of solar cost (in log base 2)") +
   scale_colour_manual(values = c("orange", "black")) 


```


# Task 3

Loading the updated dataset from the **California Solar Initative**, the CSI Working Dataset.

```{r, echo = TRUE, include = TRUE, warning = FALSE}
updated_df <- read_csv("lab7_data/WorkingDataSet_5-28-2020.csv")
```

##  Goodness-of-fit model
```{r, echo = TRUE, include = TRUE, warning = FALSE}
interval = predict(learning_mod1, newdata = new_data, interval = "prediction")
new_data["log2_cost_pred"] = interval[,1]
new_data["lwr"] = interval[,2]
new_data["upr"] = interval[,3]




```

## 